|

### 2023-03-31 manuscript/manuscript_1.0.Rmd
addressing Daniel´s comments on the manuscript

---
title: |
  | Learning can mediate both badges of status and individual recognition
author:
- Andrés E. Quiñones^[Laboratorio de Biología Evolutiva de Vertebrados, Departamento
  de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia]
- C. Daniel Cadena*
- Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
  word_document: default
fig_caption: yes
fontsize: 12pt
header-includes: "```{=latex}\n  \\usepackage{float}\n  \\usepackage{unicode-math}\n
  \ \\usepackage{setspace}\\doublespacing\n  \\usepackage[backref=page]{hyperref}\n
  \ \\linespread{2}\n  \\usepackage{lineno}\n  \\linenumbers\n  \\floatplacement{figure}{H}\n
  \ \\newcommand{\\beginsupplement}{ \\setcounter{table}{0}         \\renewcommand{\\thetable}{S\\arabic{table}}\\setcounter{figure}{0}
  \\renewcommand{\\thefigure}{S\\arabic{figure}}}\n```\n"
keep_tex: yes
keywords: My keywords
linespread: 2
bibliography: ../SocialMorphCue.bib
spacing: double
style: C:\Users\andre\Zotero\styles\elife.csl
abstract: "Adaptive behavioural responses often depend on qualities of the interacting
  partner of an individual. For example, when competing for resources, an individual
  might be better off escalating fights with individuals of lower quality, while restraining
  from fighting individuals of higher quality. Communication systems involving signals
  of quality allow individuals to reduce uncertainty regarding the fighting ability
  of their partners and make more adaptive behavioral decisions. However, dishonest
  individuals can destabilize such communications systems. One open question is whether
  cognitive mechanisms, such as learning, can maintain the honesty of signals, thus,
  favoring their evolutionary stability. Here we present evolutionary simulations,
  where individuals can produce a signal proportional to their quality; and learn
  along their lifetime the best response to the signal emitted by their peers. In
  these simulations, learning on the receiver side can mediate the evolution of signals
  of quality on the sender side. When the cost of the signal is proportional to the
  quality of the sender populations are only composed of honest signalers. When the
  cost is not proportional to the quality of the signaler, the population is composed
  of both honest and dishonest signalers. We argue that learning can be a general
  cognitive mechanism playing a role in a wide range of communication systems.  \n"
---

@@

The outcome of interactions with conspecifics individuals is an important 
determinant of fitness in social animals. Irrespective of whether 
such interactions are cooperative or competitive, the actions of 
interacting individuals partly determine 
their reproductive success. However, the best action for an individual 
engaging in a social interaction might vary depending on its 
own condition, and that of individuals with 
whom it interacts. Thus, acting based on information about interacting 
partners is typically adaptive [@quinones_Negotiation_2016].
Acquiring such information, however, is far from trivial. 
In some cases, interacting partners (signallers) might be 'willing' to 
provide accurate information, but in others it might be in their 
own interest to conceal information [@johnstone_Recognition_1997], 
or to provide deceptive information [@johnstone_Badges_1993]. 
Typically, in a given context, some individuals benefit 
from broadcasting accurate information, whereas others benefit 
from concealing it. Take, for instance, an interaction between two individuals 
where one can help the other. Given some costs and benefits, 
a potential donor would be interested in helping 
related individuals. Therefore, for a relative of the donor, 
broadcasting kinship would be advantageous, whereas concealing the lack 
of kinship would be better for an unrelated individual. 
Similar scenarios may apply to a variety of aspects of social 
life such as finding mates, feeding offspring, or engaging in
dominance relationships or aggressive interactions 
[@bradbury_Principles_2011; @moller_Female_1988; @tibbetts_Socially_2004].
In dominance and aggressive interactions, a crucial piece of
information to guide individual actions is the fighting ability of interacting 
partners, which is sometimes referred to as Resource Holding Potential 
[RHP, Parker -@parker_Assessment_1974], or simply as quality. 
Responsiveness to the quality of an opponent is central 
to communication systems such as those in which individuals 
exhibit badges of status, where a signal conveys quality 
[@johnstone_Badges_1993; @rohwer_Social_1975a]. 
Such signals must be arbitrary, meaning they are not 
ontogenetically correlated with quality. 
Generally signals can be evolutionarily stable whenever low quality individuals 
pay a higher fitness cost for carrying the signal. 
So, it's no longer in the interest of low quality individuals 
to fake quality by producing a large badge. Signal costs inversely proportional
to quality con come about by the production costs of the signal, 
in which case the signal works like a handicap [@botero_Evolution_2010; @johnstone_Badges_1993;@grafen_Biological_1990; @zahavi_Mate_1975]. 
Alternatively, the costs can come from 
the aggressive reaction of receivers when the convention of the communication 
is broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. 
<!-- should this last part of the paragraph stay ?-->
<!-- An alternative system of communication is one where individuals  -->
<!-- gather and store information about each individual they interact with.  -->
<!-- This information should reliably allow each individual to react adaptively  -->
<!-- to the quality difference with their opponent. This type of  -->
<!-- assessment strategy is often referred to as individual  -->
<!-- recognition [@whitfield_Plumage_1986]. Evidently, individual recognition  -->
<!-- is restricted to  certain interaction structures, because individuals  -->
<!-- are limited by cognitive abilities in how many individuals  -->
<!-- they can recognize and track. Thus, individual recognition and  -->
<!-- badges of status have been assumed to be alternative communication  -->
<!-- systems fulfilling the same function, reducing the costs of  -->
<!-- aggression in contests over resources.  -->

@@

Continue here!!!!

A somewhat ignored component of communication systems in the context 
of aggressive interactions are cognitive aspects of the receiver module. 
Theoretical models often assume that communication systems relying on badges of 
status have reaction norms as their mechanistic underpinning 
[@botero_Evolution_2010], with 
individuals using their own quality and the opponent's badge to determine 
whether to aggressively engage in contests. Reaction norms 
allow individuals to respond to the available information without a large
cognitive burden. The computations involved in the reaction norm response do not
required a lot of memory or advanced algorithmic processes. 
This is in contrast to other information processing 
mechanisms like individual recognition. With individual recognition, 
individuals associate cues of their peers to their quality. Because these 
associations must be learnt throughout the life of individual, 
they have the usual cognitive requirements of associative learning processes. 
An alternative view of badges of status is that individuals learn 
to react to them based on their experiences [@guilford_Receiver_1991], 
which would imply that receivers must learn to associate signals with 
the fighting ability of bearers just as in systems based on individual 
recognition. In cases where badges of status vary quantitatively 
(e.g. in size or intensity), fighting ability may increase monotonically 
with attributes of the badge and would be reinforced by every interaction.
Therefore, in principle, learning in such scenarios would be faster 
than in systems involving individual 
recognition in which the association between signals and their meaning 
varies depending on the interacting partners. In any case, learning may be 
a central cognitive mechanism in both types of communication systems, 
but the role of learning in these contexts has not been thoroughly 
explored in the empirical nor theoretical literature.

@@

```{r model-struc,echo=FALSE,warning=FALSE,fig.width=8,fig.align='center',fig.height=7,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions. In A, the reaction norm determines the badge size. Red shows an informative reaction norm, while the blue shows a uninformative reaction norm. B individuals perceiving the signal have a behavioral reaction norm that determines their probability of retreating (black line). The reaction norm arises from generalizing the information from the feature weights (black dots). Feature weaights are values associated with a particular badge size and are updated as individuals interact with each other. The generalization of the information associated with each feature weight is represented by the grey line and its axis, which shows the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center. The learning process moves the feature weights (black dots) up or down depending on whether that leads to an increase in the estimated reward. In C and D, effect of learning on the receiver strategy. Receivers in C face signallers with uninformative reaction norms (blue line in A), while in D they face signallers with informative reaction norms (red line in A). Colour scale in C and D indicates the quality of the individual"}

@@

```{r handicap,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap="Evolution of the badge size as a handicap mediated by learning. Top panels (A.1 and A.2) show the evolutionary dynamics of the sender code. On the left (A.1), changes in the distribution of values for the intercept of the reaction norm ($\\alpha$); on the right (A.2) changes in the distribution of the slope ($\\beta$). Darker areas of the background correspond to values with high frequency, while dashed lines in both panels show the mean of the distribution. X axis in the evolutionary dynamics are given in thousands of generations. Grey lines show the generation time corresponding to the panels below portraing the sender (B) and receiver code(C). In the bottom panels (C1-4), the learned reaction norms correspond to individuals in the population after the interaction round. Color scale indicates quality just as in Fig. \\ref{fig:model-struc}." ,warning=FALSE}

@@

```{r branching,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap= "The evolution of cheap signals. Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig. \\ref{fig:handicap}. The top panels (A) show changes in the distribution of values for $\\alpha$ (A.1) and $\\beta$ (A.2) along evolutionary time. Panels below correspond to snapshots of the sender (B) and receiver codes (C). Generation time of the snapshots are indicated by the grey lines in the top panels."}

@@

```{r honestDiv,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap= "The evolution of diverse costly signals. Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig. \\ref{fig:handicap}. The middle panels show changes in the distribution of values for $\\alpha$ and $\\beta$ along evolutionary time. Panels above and below correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels."}

@@

```{r startCond, echo=FALSE, out.width='100%',fig.align='center',fig.show='hold',fig.cap="The peaceful, the clever and the aggressive. Distribution of values of the intercept $\\alpha$ and slope $\\beta$ for individuals at the end of the evolutionary simulations. The panels show the three different initial conditions for the behaviour of individuals: peaceful (blue), clever (green) and aggressive (red), see maintext for details. Colours indicate the replicate simulation. In the inset, the resulting reaction norms corresponding to the intecept and slope values for one of the replicates, which replicate is shown is indicated by the line colour."}

>>>>

---
title: |
  | The role of learning in status signalling: a modeling approach
author:
- Andrés E. Quiñones^[Laboratorio de Biología Evolutiva de Vertebrados, Departamento
  de Ciencias Biológicas, Universidad de los Andes, Bogotá, Colombia]
- C. Daniel Cadena*
- Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  word_document: default
  # pdf_document:
  #   latex_engine: xelatex
fig_caption: yes
fontsize: 12pt
header-includes: "```{=latex}\n  \\usepackage{float}\n  \\usepackage{unicode-math}\n
  \ \\usepackage{setspace}\\doublespacing\n  \\usepackage[backref=page]{hyperref}\n
  \ \\linespread{2}\n  \\usepackage{lineno}\n  \\linenumbers\n  \\floatplacement{figure}{H}\n
  \ \\newcommand{\\beginsupplement}{ \\setcounter{table}{0}         \\renewcommand{\\thetable}{S\\arabic{table}}\\setcounter{figure}{0}
  \\renewcommand{\\thefigure}{S\\arabic{figure}}}\n```\n"
keep_tex: yes
keywords: My keywords
linespread: 2
bibliography: ../SocialMorphCue.bib
spacing: double
style: C:\Users\andre\Zotero\styles\elife.csl
abstract: "Adaptive behavioural responses often depend on qualities of the interacting
  partner of an individual. For example, when competing for resources, an individual
  might be better off escalating fights with individuals of lower quality, while restraining
  from fighting individuals of higher quality. Communication systems involving signals
  of quality allow individuals to reduce uncertainty regarding the fighting ability
  of their partners and make more adaptive behavioral decisions. However, dishonest
  individuals can destabilize such communications systems. One open question is whether
  cognitive mechanisms, such as learning, can maintain the honesty of signals, thus,
  favoring their evolutionary stability. Here we present evolutionary simulations,
  where individuals can produce a signal proportional to their quality; and learn
  along their lifetime the best response to the signal emitted by their peers. In
  these simulations, learning on the receiver side can mediate the evolution of signals
  of quality on the sender side. When the cost of the signal is proportional to the
  quality of the sender populations are only composed of honest signalers. When the
  cost is not proportional to the quality of the signaler, the population is composed
  of both honest and dishonest signalers. We argue that learning can be a general
  cognitive mechanism playing a role in a wide range of communication systems.  \n"
---

@@

The outcome of interactions with conspecifics individuals is an important 
determinant of fitness in social animals. Irrespective of whether 
such interactions are cooperative or competitive, the actions of 
interacting individuals partly determine 
their reproductive success. However, the best action for an individual 
engaging in a social interaction might vary depending on its 
own condition, and that of individuals with 
whom it interacts. Thus, acting based on information about interacting 
partners is typically adaptive [@quinones_Negotiation_2016], yet
acquiring such information is far from trivial. 
In some cases, interacting partners (signallers) might be 'willing' to 
provide accurate information, but in others it might be in their 
own interest to conceal information [@johnstone_Recognition_1997], 
or to provide deceptive information [@johnstone_Badges_1993]. 
Typically, in a given context, some individuals benefit 
from broadcasting accurate information, whereas others benefit 
from concealing it. Take, for instance, an interaction between two individuals 
where one can help the other. Given some costs and benefits, 
a potential donor is likely to be interested in helping 
related individuals. Therefore, for a relative of the donor, 
broadcasting kinship would be advantageous, whereas concealing the lack 
of kinship would be better for an unrelated individual. 
Similar scenarios may apply to a variety of aspects of social 
life such as finding mates, feeding offspring, or engaging in
dominance relationships or aggressive interactions 
[@bradbury_Principles_2011; @moller_Female_1988; @tibbetts_Socially_2004].

In dominance and aggressive interactions, a crucial piece of
information to guide individual actions is the fighting ability of interacting 
partners, which is sometimes referred to as Resource Holding Potential 
[RHP, Parker -@parker_Assessment_1974], or simply as quality. 
Responsiveness to the quality of an opponent is central 
to communication systems such as those in which individuals 
exhibit badges of status, where a signal conveys quality 
[@johnstone_Badges_1993; @rohwer_Social_1975a]. 
Such signals must be arbitrary, meaning the signals could potentially be 
produced by high and low quality individuals alike. 
Generally signals can be evolutionarily stable whenever low quality individuals 
pay a higher fitness cost for carrying the signal such that it is not 
in the interest of low quality individuals 
to fake quality by producing a large badge. Signal costs inversely proportional
to quality can come about as a result of the production costs of the signal, 
in which case the signal works like a handicap [@botero_Evolution_2010; @johnstone_Badges_1993;@grafen_Biological_1990; @zahavi_Mate_1975]. 
Alternatively, the costs can come from 
the aggressive reaction of receivers when the convention of the communication 
is broken [@enquist_Signaling_2010;@tibbetts_Socially_2004]. 
<!-- should this last part of the paragraph stay ?-->
<!-- An alternative system of communication is one where individuals  -->
<!-- gather and store information about each individual they interact with.  -->
<!-- This information should reliably allow each individual to react adaptively  -->
<!-- to the quality difference with their opponent. This type of  -->
<!-- assessment strategy is often referred to as individual  -->
<!-- recognition [@whitfield_Plumage_1986]. Evidently, individual recognition  -->
<!-- is restricted to  certain interaction structures, because individuals  -->
<!-- are limited by cognitive abilities in how many individuals  -->
<!-- they can recognize and track. Thus, individual recognition and  -->
<!-- badges of status have been assumed to be alternative communication  -->
<!-- systems fulfilling the same function, reducing the costs of  -->
<!-- aggression in contests over resources.  -->

@@

A somewhat ignored component of communication systems in the context 
of aggressive interactions are cognitive aspects of the receiver module. 
Theoretical models often assume that communication systems relying on badges of 
status have reaction norms as their mechanistic underpinning 
[@botero_Evolution_2010], with 
individuals using their own quality and the badges of opponents to determine 
whether to aggressively engage in contests. Reaction norms 
allow individuals to respond to the available information without a large
cognitive burden because the computations involved in the reaction 
norm response do not require much memory nor advanced algorithmic processes. 
This contrasts with other information-processing 
mechanisms like individual recognition. With individual recognition, 
where individuals associate cues of their peers to their quality. Because these 
associations must be learnt throughout the life of individuals, 
they have the usual cognitive requirements of associative learning processes. 
An alternative view of badges of status is that individuals learn 
to react to them based on their experiences [@guilford_Receiver_1991], 
which would imply that receivers must learn to associate signals with 
the fighting ability of bearers just as in systems based on individual 
recognition. In cases where badges of status vary quantitatively 
(e.g. in size or intensity), fighting ability may increase monotonically 
with attributes of the badge and would be reinforced by every interaction.
Therefore, in principle, learning in such scenarios would be faster 
than in systems involving individual 
recognition in which the association between signals and their meaning 
varies depending on the interacting partners. In any case, learning may be 
a central cognitive mechanism in both types of communication systems, 
but the role of learning in these contexts has not been thoroughly 
explored in the empirical nor theoretical literature.

<!-- Continuar aquí en la revisión -->

@@

```{r model-struc,echo=FALSE,warning=FALSE,fig.width=8,fig.align='center',fig.height=7,fig.show='hold',fig.cap= "Model of communication in the context of aggressive interactions. In A, the reaction norm determines the badge size. Red shows an informative reaction norm, while the blue shows a uninformative reaction norm. B individuals perceiving the signal have a behavioral reaction norm that determines their probability of retreating (black line). The reaction norm arises from generalizing the information from the feature weights (black dots). Feature weaights are values associated with a particular badge size and are updated as individuals interact with each other. The generalization of the information associated with each feature weight is represented by the grey line and its axis, which shows the response triggered by the fourth feature diminishes as the value evaluated is further from the feature center. The learning process moves the feature weights (black dots) up or down depending on whether that leads to an increase in the estimated reward. In C and D, effect of learning on the receiver strategy. Receivers in C face signallers with uninformative reaction norms (blue line in A). The uninformative reaction norm yields badge size of 0.5, thus the behavioural reaction norm in C is only updated around that value.D Receivers face signallers with informative reaction norms (red line in A). Accordingly, receivers develop a threashold-like reaction norm where the decision to retreat increases with the badge size of the interacting partner. Colour scale in C and D indicates the quality of the individual. "}

@@

```{r handicap,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap="Evolution of the badge size as a handicap mediated by learning. Top panels (A.1 and A.2) show the evolutionary dynamics of the sender code. On the left (A.1), changes in the distribution of values for the intercept of the reaction norm ($\\alpha$); on the right (A.2) changes in the distribution of the slope ($\\beta$). Darker areas of the background correspond to values with high frequency, while dashed lines in both panels show the mean of the distribution. X axis in the evolutionary dynamics are given in thousands of generations. Grey lines show the generation time corresponding to the panels below portraing the sender (B) and receiver code(C). In the bottom panels (C1-4), the learned reaction norms correspond to individuals in the population after the interaction round. Color scale indicates quality just as in Fig. \\ref{fig:model-struc}. On the first half of the simulation (B.1 and B.2) the badge size evolves to its minimum value. In the second half, the value of the slope takes positive values and reaction norms evolve to produce an honest signal of quality. Correspondingly, individuals learn to react to badge size by increasing the probability of retreat with increasing badge sizes"  ,warning=FALSE}

@@

```{r branching,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap= "The evolution of cheap signals. Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig. \\ref{fig:handicap}. The top panels (A) show changes in the distribution of values for $\\alpha$ (A.1) and $\\beta$ (A.2) along evolutionary time. Panels below correspond to snapshots of the sender (B) and receiver codes (C). Generation time of the snapshots are indicated by the grey lines in the top panels. The unimodal distribution with which $\\alpha$ (A.1) starts the simulation, quickly turns into a bimondal distribution, and at about 10000 generations it splits further into three modes. These three peaks correspond to three type of reaction norms in panels B.1-3."}

@@

```{r honestDiv,echo=FALSE,out.width='100%',fig.height=6,fig.align='center',fig.show='hold',fig.cap= "The evolution of costly signals. Portrait of the evolutionary dynamics of the sender code with snapsots of both sender and receiver codes just as in fig. \\ref{fig:handicap}. The middle panels show changes in the distribution of values for $\\alpha$ and $\\beta$ along evolutionary time. Panels above and below correspond to snapshots of the sender and receiver codes, respectively, generation time of the snapshots are indicated by the grey lines in the middle panels. Similarly to the evolutionary dynamics in Fig. \\ref{fig:handicap}, first the badge size evolves towards its minimum value. Then, the slope evolves positive values determining an increasing reaction norm. In contrast to \\ref{fig:handicap}, the normal distribution splits up into two modes. This translates into two types of reaction norms assorting in the population where both code for a positive relation between quality and badge size."}

@@

```{r startCond, echo=FALSE, out.width='100%',fig.align='center',fig.show='hold',fig.cap="The peaceful, the clever and the aggressive. Distribution of values of the intercept $\\alpha$ and slope $\\beta$ for individuals at the end of the evolutionary simulations. The panels show the three different initial conditions for the behaviour of individuals: peaceful, \"clever\" and aggressive, see maintext for details. Colours indicate the replicate simulation. In the inset, the resulting reaction norms corresponding to the intecept and slope values for one of the replicates, which replicate is shown is indicated by the line colour. The simulations where individual disply a \"clever\" and agressive behaviour, before the learning process, show all the points of a single replicate cluster together, while in simulation where individuals start being peaceful al replicates show distinct groups of individuals spread throughout the x and y axis. Only in this condition there is a frequency-dependent process that favour diversity in the reaction norm."}
